\documentclass[twoside]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumerate}% http://ctan.org/pkg/enumerate
\usepackage{pgfplots}
\usepackage{multicol}
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage{fullpage}
\usepackage{pdflscape}
\usepackage{setspace}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{tikzmark}
\usepackage{color}
\usepackage{pgfplots}
\usepackage{bm}
\usepackage{tikz-cd}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{fillbetween}
\usepackage[toc,page]{appendix}
\usepackage[shortlabels]{enumitem}\usepackage{draftwatermark}

\usetikzlibrary{arrows.meta,
                calc, chains,
                patterns,
                patterns.meta,
                fit,
                quotes,
                positioning,
                shapes.geometric,
                shapes.misc,
                automata
                }
%Flowchart Tikzset
\tikzset{
   recbox/.style = {
         rectangle,
         draw, 
         align = center, 
         text badly centered,
         inner sep = 6 pt,
         font=\footnotesize,
         line width = 0.3mm,
      },
      circlebox/.style = {
         rounded rectangle,
         draw, 
         align = center, 
         text badly centered,
         inner sep = 7 pt,
         font=\large,
         line width = 0.5mm,
      },
      roundbox/.style = {
         rectangle,
         draw, 
         align = center, 
         rounded corners,
         text badly centered,
         inner sep = 6 pt,
         font=\large,
         line width = 0.5mm,
      },
     box1/.style = {
         rectangle,
         draw, 
         align = center, 
         text badly centered,
         inner sep = 6 pt,
         font=\large,
         line width = 0.5mm,
         minimum width = 30mm,
         minimum height = 7mm,
      },
    papLine/.style = {
         draw,
         -stealth,
         font=\ttfamily,
         line width = 0.5mm,
      },
      }
%end flowcahrt tikz set   

%Colors 
\definecolor{darkgreen}{rgb}{0.01, 0.75, 0.24}
\definecolor{shaded_green}{RGB}{151, 194, 157}

\newcommand{\quotes}[1]{``#1''}

\theoremstyle{plain}% default
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem*{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{exercise}[example]{Exercise}

\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}

\begin{document}
\parindent=0in
\parskip=12pt

\SetWatermarkText{Draft}
\SetWatermarkScale{5}

\title{
  Relative Probability on Finite Sample Spaces \\
  \large{
    Topological Addendum
  }
}

\author{Max Sklar\\ Local Maximum Labs \\ DATE HERE}
\date{}

\maketitle
\thispagestyle{empty}

\begin{abstract}
This is an incomplete draft/outline of an upcoming paper. Please do not share
\end{abstract}

\tableofcontents
\newpage

\section{The Relative Probability Approach}
\label{section:new_relative_prob}

In section \ref{section:standard_relative_prob}, the relative probability function was derived from the absolute probability function. Here in section \ref{section:new_relative_prob}, we start with the relative probability function as the fundamental object of study.

\subsection{Fundamental Axioms}

Consider a relative probability function \(P\) that acts on outcomes in \(\Omega\).

\begin{definition}
\label{def:fundamental_laws}
Let \(\Omega\) be the set of outcomes, and \(P: \Omega \times \Omega \rightarrow \mathbb{M}^*\) be a function acting on two outcomes to produce a magnitude-wildcard. \(P\) is a \textit{relative probability function on the outcomes of \(\Omega\)} if it obeys the \textit{3 fundamental axioms of relative probability}:

\begin{enumerate}[(i)]
\item The \textit{identity axiom}: \(P(h, h) = 1\)
\item The \textit{inverse axiom}: \(P(h_1, h_2) = P(h_2, h_1)^{-1}\)
\item The \textit{composition axiom}: \(P(h_1, h_3) :\cong P(h_1, h_2) \cdot P(h_2, h_3)\)
\end{enumerate}

\end{definition}

\(P(h_1, h_2)\) can be read as the probability of \(h_1\) relative to \(h_2\). Outcomes \(h_1\) and \(h_2\) are \textit{comparable} if \(P(h_1, h_2) \neq \ast\).

Let us pause for a moment to discuss how these axioms were chosen. The star of the show is the composition axiom which succinctly encodes how relative probability works. If \(A\) is twice as likely as \(B\), and \(B\) is 3 times as likely as \(C\), then \(A\) had better be 6 times as likely as \(C\). If not, these relative probability assignments would have no meaning; they would just be numerical assignments without rhyme or reason\footnote{Many of our political and economic forecasts come in this form.}.

The composition axiom is enough to show that the identity axiom works most of the time. For example, if \(h_1\) is comparable to any other outcome \(h_2\) then through composition we get \(P(h_1, h_2) :\cong P(h_1, h_1) \cdot P(h_1, h_2)\). So long as \(P(h_1, h_2)\) isn't \(0\), \(\infty\), or \(\ast\), then we would have to conclude \(P(h_1, h_1) = 1\).

But that doesn't get us all the way there! We can still construct scenarios where \(P(h, h) = \ast\). The self-comparisons in an outcome space should not be able to contain any information where there is a choice of values. Hence, the neccesity of the identity axiom.

Composition and identity can actually be combined into a single axiom about composition paths. It's a bit unweildy for the mathematical proofs, but nevertheless interesting.

\begin{proposition}[Path Composition]
Given a non-empty list of \(N\) outcomes \(h_0, h_1, h_2, ..., h_{N-1}\), \[P(h_0, h_{N-1}) :\cong \prod_{k=0}^{N-2} P(h_k, h_{k+1}) \]
\end{proposition}

In this case, \(P(h_0, h_0)\) would be matched by the empty product, which is 1.

The inverse axiom is nearly redundant as well. Since \(P(h_0, h_0) \cong P(h_0, h_1) \cdot P(h_1, h_0)\), the terms in the constraint look like they must be inverses! But without stating the axiom explicitly, there could be a case where \(P(h_0, h_1)\) is some non-wildcard magnitude like 2 but \(P(h_1, h_0)\) is \(\ast\). This shouldn't be allowed because \(\ast\) represents a lack of knowledge about a value, and we consider \(P(h_1, h_0)\) and \(P(h_1, h_0)\) to be the same piece of information but in reverse.

\subsection{Examples}

Now that the definition of relative probability is squared away, we can construct a library of examples for common RPFs that will serve as building blocks to tackling common problems.

\begin{definition}
\label{def:uniform_rpf}
The \textit{uniform} RPF can be constructed from any number of outcomes where each are considered equally likely. \(P(h_1, h_2) = 1\) for every pair of outcomes.
\end{definition}

\begin{definition}
\label{def:uncomparable_rpf}
The \textit{uncomparable} RPF has \(P(h_1, h_2) = \ast\) for every pair of outcomes. It is as if the subjective probability agent gave up or the Bayesian model was fed corrupt data.
\end{definition}

\begin{definition}
A \textit{certain} RPF contains a single outcome that has infinite probability relative to all other outcomes. Let \(h_C\) be the certain outcome with \(h_C \neq h\). Then \(P(h_C, h) = \infty\). The relative probability of the other \(K-1\) outcomes could be anything.
\end{definition}

\begin{definition}
\label{def:empty_rpf}
The \textit{empty} RPF has no outcomes \(K = 0\), and therefore the function \(P\) has no valid inputs.
\end{definition}

It is surprising that there is still an RPF with \(\Omega = \varnothing\). This is not the case for absolute distributions where such a function does not exist (because with no outcomes, they cannot sum to 1).

\begin{definition}
\label{def:unit_rpf}
The \textit{unit} RPF has a single outcome where \(K = 1\) and \(\Omega = h\). There is only one such RPF where \(P(h, h) = 1\).
\end{definition}

The unit RPF is a special case of the uniform RPF and the certain RPF. This matches the absolute case where the probability of the single outcome must be 1.

\begin{definition}
\label{def:finite_geometric_rpf}
Let \(P\) be an RPF with K outcomes labeled \((h_0, h_1, ..., h_{K-1})\). \(P\) is a \textit{finite geometric} RPF with ratio \(r\) if the relative probabilities of each outcome with its neighbor is always \(r\). In other words, for all \(i \in (0, 1, ..., K-2)\),
\[P(h_{i+1}, h_i) = r\]
When \(r\) is 0 or \(\infty\), we can call this the \textit{limit finite geometric} RPF.
\end{definition}

Finally, to include an example that is both common and has powerful applications, there is a relative version of the Binomial distribution.

\begin{definition}
\label{def:binomial_rpf}
A \textit{binomial distribution} has a sample size we can call \(n\), and a probability of success \(p\). The RPF is uses \(\Omega = \{0, 1, 2, ..., n\}\), and thus \(K = n + 1\). It is given as follows:
\[P(h_1, h_2) = \frac{h_2!(n-h_2)!}{h_1!(n-h_1)!}\left(\frac{p}{1-p}\right)^{h_1 - h_2}\]
\end{definition}

\section{New Concepts for Relative Probability}

We have successfully defined the relative probability in section \ref{section:new_relative_prob} with fundamental axioms and have constructed some examples. Because new situations arise that do not occur in the Kolmogorov model, we need to define some new vocabulary.

Fortunately, the absolute probability function is a special case of relative probability through definition \ref{def:ratio}, defined by \(P(h_1, h_2) = \frac{P(h_1)}{P(h_2)}\). This ratio can be adjusted to follow all of the fundamental axioms, notably composition from theorem \ref{thm:abs_composition}.

Figure \ref{fig:flow_chart} gives us a roadmap of these new concepts and their relationship to each other.

\begin{figure}
\begin{tikzpicture}    %Flowchart
[node distance = 1.3in] %Node distance

%node0 for Relative Probability Functions
\node (node0) [roundbox] {Relative Probability Function \\ Definition \ref{def:fundamental_laws}} ;

%node1 Anchored RPF
\node (node1) [circlebox, below of=node0] {Anchored RPF\\ Definition \ref{def:anchored_rpf}};

%lnode1 Empty RPF
\node (lnode1) [recbox, fill=shaded_green,below left=0.5 and 2.2 of node0] {Empty RPF\\ Definition \ref{def:empty_rpf}};

%lnode2 Uncomparable RPF
\node (lnode2) [recbox, fill=shaded_green,below left=2 and 0.5 of node0] {Uncomparable RPF\\ Definition \ref{def:uncomparable_rpf}};

%node2 Non-empty Totally Comparable RPF
\node (node2) [roundbox, below right= and 2 of node1] {Non-empty\\ Totally Comparable RPF\\ Definition \ref{def:totally_comparable}};

%bnode2 Finite Limit Geometric Distribution, \(K>2\)
\node (bnode2) [recbox, fill=shaded_green, below right=1 and -3.5 of node2] {Finite Limit \\Geometric Distribution,\\ \(K>2\)\\ Definition \ref{def:finite_geometric_rpf}};

%node3 Absolute Probability\\ Function
\node (node3) [roundbox, below left= and 2 of node1] {Absolute Probability\\ Function\\ Definition \ref{def:categorical_abs}};

%node4 Union
\node (node4) [circlebox, below right=and 2.8 of node3] {Union};

%node5 Facet Probability\\ Functions\\(single impossible event)
\node (node5) [box1, below left=and -0.5 of node4,yshift=-0.38in] {Facet Probability\\ RPF\\(single impossible event)\\};

%node6 Mutually Possible\\ RPFs
\node (node6) [box1, right=-0.07 of node5] {Totally Mutually Possible\\ RPF\\ Definition \ref{def:totally_mutually_possible}\\};

%rnode4 (Non-Empty)\\ Uniform
\node (rnode4) [recbox, fill=shaded_green, below of=node6,xshift=-1.0in,yshift=-0.0in] {(Non-Empty)\\ Uniform\\ Definition \ref{def:uniform_rpf}};

%rnode3 Unit\\ Distribution
\node (rnode3) [recbox, fill=shaded_green, below of=node6,yshift=-0.1in] {Unit\\ Distribution\\ Definition \ref{def:unit_rpf}};

%rnode2 Finite Geometric\\Distribution
\node (rnode2) [recbox, fill=shaded_green, below of=node6,xshift=1.1in,yshift=-0.0in] {Finite Geometric\\Distribution\\ Definition \ref{def:finite_geometric_rpf}};

%rnode1 Binormal\\Distribution
\node (rnode1) [recbox, fill=shaded_green, right=1 of node6] {Binormal\\Distribution\\ Definition \ref{def:binomial_rpf}};

%lnode3 Limit Finite\\Geometry\\ with \(K=2\)
\node (lnode3) [recbox, fill=shaded_green, below of=node5,yshift=-0.0in,xshift=-0.5in] {Limit Finite\\Geometric\\ with \(K=2\)};

%Arrow paths for all nodes
\path[papLine] (rnode1) -- (node6);

\path[papLine] (rnode2) --+ (node6);

\path[papLine] (rnode3) -- (node6);

%\path[papLine] (rnode4) --+(0,0.56in)--+(1in,0.56in)-- +(1in,0.87in); (node6);
\path[papLine] (rnode4) --+ (node6);

\path[papLine] (lnode1) --+ (node0);

\path[papLine] (lnode2) --+ (node0);

\path[papLine] (node1) -- (node0);

\path[papLine] (node2) -- (node1) node[pos=0.4, above, sloped]{Lemma \ref{lemma:totally_comp_anchored}};

\path[papLine] (node3) -- (node1) node[pos=0.5, above, sloped]{Theorem \ref{thm:absolute_anchored}};

\path[papLine] (node4) -- (node3);

\path[papLine] (node4) -- (node2);

\draw[-stealth,transform canvas={xshift=1mm},line width=0.5mm] (lnode3) -- +(0,0.86in);

\draw[-stealth,transform canvas={xshift=24mm},line width=0.5mm] (node5) --+(0,1.22in) node[left,pos=0.5]{Theorem \ref{thm:abs_totally_comparable}};

\path[papLine,dashed] (node1) |- (node3) node[above,pos=0.75]{Matched By} node[below,pos=0.75]{Theorem \ref{thm:absolute_prob_formula}};

\path[papLine] ([xshift=0.6in]bnode2) -- (node2);
\end{tikzpicture}
\caption{This is our roadmap for all of the sub-types of relative probability functions and their relationship to one another. }
\label{fig:flow_chart}
\end{figure}

\subsection{Matching and Comparability}

\begin{definition}
\label{def:totally_comparable}
A relative probability function is \textit{totally comparable} if every pair of outcomes are comparable.
\end{definition}

\begin{theorem}
\label{thm:abs_totally_comparable}
An absolute probability function is totally comparable if and only if \(P(h) = 0\) for at most one outcome.
\end{theorem}

\begin{proof}
Let P be an \textbf{absolute} probability function, with \(h_1\) and \(h_2\) being two outcomes. If \(P(h_1) = P(h_2) = 0\), then \(P(h_1, h_2) = \frac{0}{0} = \ast\). If only outcome \(h_1\) is assigned 0, then \(P(h_1, h_1) = 1\), \(P(h_1, h_2) = 0\), and \(P(h_2, h_1) = \infty\). Any other pairing that does not involve \(h_1\) will be the quotient of two positive numbers, and thus also comparable.
\end{proof}

\begin{definition}
\label{def:anchored_rpf}
An \(anchored\) RPF has at least 1 outcome whose probability relative to every other outcome is greater than zero. We call this outcome an \(anchor\).
\end{definition}

Anchor outcomes are those outcomes that have a non-zero absolute probability. The anchoring of a distribution ensures that it is well behaved.

\begin{theorem}
\label{thm:absolute_anchored}
All absolute probability distributions are anchored.
\end{theorem}

\begin{proof}
Let P be an absolute probability distribution on \(\Omega\). Because \(\sum_{h \in \Omega} P(h) = 1\), there must be at least one \(h\) such that \(P(h) > 0\).  Therefore, for any comparison outcome \(h'\), \(P(h, h') > 0\)
\end{proof}

\begin{lemma}
\label{lemma:totally_comp_anchored}
Every non-empty, totally comparable RPF is anchored.
\end{lemma}

\begin{proof}
Let \(P\) be non-empty and totally comparable RPF. Assume the opposite - that is for every outcome \(h\), there exists another outcome \(h'\) such that P(h, h') = 0.

A function \(f: \Omega \rightarrow \Omega\) can be created so that for every \(h\), \(P(h, f(h)) = 0\).

Let \(f^n\) be the function \(f\) applied n times. Then \(P(h, f^n(h)) = 0\) for all n greater than 0. This is by induction because the case of \(n = 1\) was assumed above, and for for inductive step
\[P(h, f^{n+1}(h)) :\cong P(h, f^n(h)) \cdot P(f^n(h), f(f^n(h)) = 0 \cdot 0 = 0\]

Because \(\Omega\) is finite, repeated applications of \(f\) on \(h\) must evenually return to an outcome that has already been visited. In more rigorous terms, there exists an \(N\) such that \(f^N(h) = f^i(h)\) for some \(i < N\).

But this is a contradiction because \(P(f^i(h), f^N(h))\) should equal 0 by the argument above, but 1 by the identity axiom.
\end{proof}

A totally comparable RPF contains the maximum amount of information about the relative probability of two outcomes. Some RPFs have less information but are nevertheless consistent with RPFs that have more. The following definition encapulates this relationship.

\begin{definition}
Let \(P_1\) and \(P_2\) be relative probability functions. \(P_1\) is matched by \(P_2\) if and only if all of relative probabilities of \(P_1\) are matched by those of \(P_2\). For all outcomes \(h_1\) and \(h_2\),
\[P_1(h_1, h_2) :\cong P_2(h_1, h_2)\]
\end{definition}

\begin{theorem}
\label{thm:absolute_prob_formula}
Every anchored RPF is matched by an absolute probability function, given by the following equation where \(a\) is an anchor outcome.
\[P(h) = \frac{P(h, a)}{\sum_{h' \in \Omega}P(h', a)}\]
\end{theorem}

\begin{proof}
We need to show that \(P(h)\) is a valid absolute probability function, and that it matches the original RPF.

Because \(a\) is an anchor element, we know that \(P(h', a) < \infty\). This means that the sum \(\sum_{h' \in \Omega}P(h', a) < \infty\). It is also non-zero, because included in that sum is \(P(a, a) = 1\). The numerator \(P(h, a)\) is also a magnitude \(< \infty\). Therefore, this formula yields \(P(h) \notin \{\infty, \ast\}\).

We next check that the values of \(P(h)\) sum to 1 as follows:
\[\sum_{h \in \Omega}P(h) = \sum_{h \in \Omega} \frac{P(h, a)}{\sum_{h' \in \Omega}P(h', a)} = \frac{\sum_{h \in \Omega}P(h, a)}{\sum_{h' \in \Omega}P(h', a)} = 1\]

Cancellation of these equal sums is justified because we have argued above that they cannot be \(0\) or \(\infty\).

Therefore, \(P(h)\) is a valid absolute probability function. The relative probability function is shows to be matched by it though calculation:
\begin{equation}
P(h_1, h_2) :\cong P(h_1, a) \cdot P(a, h_2) = \frac{P(h_1, a)}{\sum_{h' \in \Omega}P(h', a)} \div \frac{P(h_2, a)}{\sum_{h' \in \Omega}P(h', a)} = \frac{P(h_1)}{P(h_2)}
\end{equation}
\end{proof}

\subsection{Mutual Possibility}

With relative probability, it is important to know not only which outcomes are comparable, but whether their relative probability is positive and finite (that is, not equalt to 0, \(\infty\), or \(\ast\)). We start with a few definitions.

\begin{definition}
Outcome \(h_1\) is \textit{impossible} with respect to \(h_2\) if \(P(h_1, h_2) = 0\). Outcome \(h_1\) is \textit{possible} with respect to \(h_2\) if \(P(h_1, h_2) > 0\).
\end{definition}

\begin{definition}
Outcomes \(h_1\) and \(h_2\) are \textit{mutually possible} if they are comparable and \(0 < P(h_1, h_2) < \infty\).
\end{definition}

\begin{theorem}
The relationship of mutually possible events is an \textit{equivalence relation}, being reflexive, symmetric and transitive.
\end{theorem}

\begin{proof}
For reflexive, \(P(h_1, h_1) = 1\) by the identity axiom.

For symmetric, \(P(h_1, h_2) = P(h_2, h_1)^{-1}\), which means that each can be in \(\{0, \infty, \ast\}\) if and only if the other is as well.

For transitive, use the composition axiom which states that \(P(h_1, h_3) :\cong P(h_1, h_2) \cdot P(h_2, h_3)\). If the last 2 values are positive and finite, then their product is also positive and finite.
\end{proof}

\begin{definition}
\label{def:totally_mutually_possible}
A relative probability function is called \textit{totally mutually possible} if all of its outcomes\footnote{Note that this one of the few definitions that cannot later be upgraded from outcomes to events. The empty event \(e = \{\}\) for example will be impossible with respect to any outcome by theorem \ref{thm:empty_event_impossible}.} are mutually possible.
\end{definition}

In a totally mutually possible RPF, every outcome is an anchor because it will have a \(>0\) probability compared to every other outcome.

It is helpful to make diagrams of possibility and impossibility through a \textit{directed graph}. In these graphs, each outcome is represented by a point, and an arrow from A to B means that B is possible with respect to A. A bidirectional arrow means that A and B are mutually possible. Totally mutually possible RPFs have a simple diagram where all the outcomes are completely connected as in figure \ref{fig:mutually_possible_rpf}.

\begin{figure}[h]
\centering
\resizebox{0.2\textwidth}{!}{
\begin{tikzpicture} %Totally Possible
\node(t)[circle,draw,inner sep=4pt,outer sep=1pt,line width=0.2mm,fill=black] at (0,2) {}; 
\node(ml)[circle,draw,inner sep=4pt,outer sep=1pt,line width=0.2mm,fill=black] at (-2,0.5){}; 
\node(mr)[circle,draw,inner sep=4pt,outer sep=1pt,line width=0.2mm,fill=black] at (2,0.5) {}; 
\node(bl)[circle,draw,inner sep=4pt,outer sep=1pt,line width=0.2mm,fill=black] at (-1.3,-1.8){}; 
\node(br)[circle,draw,inner sep=4pt,outer sep=1pt,line width=0.2mm,fill=black] at (1.3,-1.8){}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(bl);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(br);

\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(br);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(bl);

\path[stealth-stealth,draw,line width=0.3mm] (mr)--(br);
\path[stealth-stealth,draw,line width=0.3mm] (bl)--(br);

\end{tikzpicture}
}
\caption{A totally mutually possible RFP has - unsurprisingly - a complete graph of mutually possibility.}
\label{fig:mutually_possible_rpf}
\end{figure}

\begin{theorem}
A non-empty totally mutually possible RPF is equal to an absolute probability function.
\end{theorem}

\begin{proof}
If \(P\) is totally mutually possible, then all of its outcomes are anchors. Therefore, we can use theorem \ref{thm:absolute_prob_formula} to find a matching absolute probability function
\[P(h) = \frac{P(h, a)}{\sum_{h' \in \Omega}P(h', a)}\]

Because every element of \(\Omega\) is an anchor, we can let \(a = h\) and get
\[P(h) = \frac{P(h, h)}{\sum_{h' \in \Omega}P(h', h)}=\frac{1}{\sum_{h' \in \Omega}P(h', h)}\]

Theorem \ref{thm:absolute_prob_formula} states that \(P(h_1, h_2) :\cong \frac{P(h_1)}{P(h_2)}\), but since the constraint is never \(\ast\), they must be equal.
\end{proof}

\subsection{Possibility Classes}

In order to analyze the general case of RPFs, we need to consider classes of mutual possibility.

\begin{theorem}
The relationship of being possible is a \textit{preorder}, being both reflexive and transitive.
\end{theorem}

\begin{proof}
It must be reflexive because \(P(h, h) = 1\). If \(P(h_1, h_2) > 0\) and \(P(h_2, h_3) > 0\) then their product is also greater than zero, and by composition, equal to \(P(h_1, h_3)\). Thus \(h_1\) is also possible with respect to \(h_3\).
\end{proof}

If we consider a possibility relationship with respect to the equivalence classes of mutually possibility, then we have a \textit{partial order}. Figure \ref{fig:anchored_rpf} is an example of outcomes grouped by mutually possible equivalence classes, with each class being impossible with respect to the ones it points to. Figure \ref{fig:anchored_rpf} is anchored while figure \ref{fig:unanchored_rpf} is not anchored.

\begin{figure}[h]
\centering
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture} %Anchored
\begin{scope}[scale=0.25]
\node(t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (0,2) {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (-2,0.5){}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (2,0.5) {}; 
\node(bl)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (-1.3,-1.8){}; 
\node(br)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (1.3,-1.8){}; 

\path[stealth-stealth,draw,line width=0.2mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(bl);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(br);

\path[stealth-stealth,draw,line width=0.2mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.2mm] (ml)--(br);
\path[stealth-stealth,draw,line width=0.2mm] (ml)--(bl);

\path[stealth-stealth,draw,line width=0.2mm] (mr)--(br);
\path[stealth-stealth,draw,line width=0.2mm] (bl)--(br);

\node[circle,fill=darkgreen, fill opacity = 0.5, draw=black,line width=0.3mm, fit=(t) (ml) (mr) (bl)(br), inner sep=-1pt] (r1) {};
\end{scope}


\begin{scope}[yshift=-0.8in,xshift=0.7in]
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node (b)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below=0.8 of t] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.35 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.35 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(t);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr) (b), inner sep=-1.8pt] (r2) {};
\end{scope}

\begin{scope}[yshift=-0.88in,xshift=-0.7in]
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.5 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.5 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr), inner sep=1pt] (r3) {};
\end{scope}

\begin{scope}[yshift=-2.1in,xshift=-0.89in]
\node(l)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black]{}; 
\node(r)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,right=0.7 of l]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (l)--(r);

\node[circle, draw=black,line width=0.3mm, fit=(l) (r), inner sep=4.7pt] (r4) {};
\end{scope}

\path[-stealth,draw,line width=0.3mm] (r2) -- (r1);
\path[-stealth,draw,line width=0.3mm] (r3) -- (r1);
\path[-stealth,draw,line width=0.3mm] (r4) -- (r3);

\end{tikzpicture}
}
\caption{A diagram of an anchored RPF with its mutually possible classes. The anchor class is the maximal class in the partial order. It is shaded.}
\label{fig:anchored_rpf}
\end{figure}

\begin{figure}[h]
\centering
\resizebox{0.5\textwidth}{!}{
\begin{tikzpicture} %Non-Anchored
\begin{scope}[scale=0.25]
\node(t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (0,2) {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (-2,0.5){}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (2,0.5) {}; 
\node(bl)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (-1.3,-1.8){}; 
\node(br)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] at (1.3,-1.8){}; 

\path[stealth-stealth,draw,line width=0.2mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(bl);
\path[stealth-stealth,draw,line width=0.2mm] (t)--(br);

\path[stealth-stealth,draw,line width=0.2mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.2mm] (ml)--(br);
\path[stealth-stealth,draw,line width=0.2mm] (ml)--(bl);

\path[stealth-stealth,draw,line width=0.2mm] (mr)--(br);
\path[stealth-stealth,draw,line width=0.2mm] (bl)--(br);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr) (bl)(br), inner sep=-1pt] (r1) {};
\end{scope}


\begin{scope}[yshift=-0.8in,xshift=0.7in]
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node (b)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below=0.8 of t] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.35 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.35 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(t);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr) (b), inner sep=-1.8pt] (r2) {};
\end{scope}

\begin{scope}[yshift=-0.88in,xshift=-0.63in]
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.5 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.5 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr), inner sep=1pt] (r3) {};
\end{scope}

\begin{scope}[yshift=0in,xshift=0.9in]
\node(l)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black]{}; 

\node[circle, draw=black,line width=0.3mm, fit=(l), inner sep=12.8pt] (r6) {};
\end{scope}

\begin{scope}[xshift=1.5in]
\node(l)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black]{}; 
\node(r)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,right=0.7 of l]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (l)--(r);

\node[circle, draw=black,line width=0.3mm, fit=(l) (r), inner sep=4.7pt] (r4) {};
\end{scope}

\begin{scope}[yshift=-1.01in,xshift=1.7in]
\node(l)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black]{}; 

\node[circle, draw=black,line width=0.3mm, fit=(l), inner sep=12.8pt] (r5) {};
\end{scope}

\path[-stealth,draw,line width=0.3mm] (r2) -- (r1);
\path[-stealth,draw,line width=0.3mm] (r3) -- (r1);
\path[-stealth,draw,line width=0.3mm] (r2) -- (r6);
\path[-stealth,draw,line width=0.3mm] (r5) -- (r4);
\end{tikzpicture}
}
\caption{This is the diagram for a single RPF that is not anchored. We cannot turn this into an absolute probability function.}
\label{fig:unanchored_rpf} 
\end{figure}

Finally, we look at totally comparable RPFs, where the graph of mutually possible components is a straight line (see figure \ref{fig:totally_comparable_rpf}).

\begin{theorem}
If an RPF is totally comparable, then the equivalance classes of mutually possible outcomes are \textit{totally ordered}. That is, each member of an equivalence class of outcomes is comparable to each member of another class with that comparison always being 0 or \(\infty\).
\end{theorem}

\begin{proof}
Let A and B be 2 distinct mutually possible equivalence classes on \(\Omega\), and let \(a \in A\) and \(b \in B\). Then \(P(a, b)\) must be either 0 or \(\infty\) because if it were in between then \(a\) and \(b\) would be in the same equivalence class, and if it were \(\ast\) then \(P\) wouldn't be totally comparable.

Let \(a' \in A\) and \(b' \in B\). Then \(0 < P(a', a) < \infty\) and \(0 < P(b, b') < \infty\) due to the definition of mutual comparability. Thus with composition we get
\[P(a', b') :\cong P(a', a) \cdot P(a, b) \cdot P(b, b') = P(a, b)\]

Therefore, all comparisons between the 2 classes will be the same, and they will either be 0 or \(\infty\).
\end{proof}

\begin{figure}[h]
\centering
\resizebox{0.1\textwidth}{!}{
\begin{tikzpicture} %Totally Comparable
\begin{scope}
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.5 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.5 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);

\node[circle, fill=darkgreen, fill opacity = 0.5, draw=black,line width=0.3mm, fit=(t) (ml) (mr), inner sep=1pt] (r1) {};
\end{scope}

\begin{scope}[yshift=-1.1in,xshift=-0.2in]
\node(l)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black]{}; 
\node(r)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,right=0.7 of l]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (l)--(r);

\node[circle, draw=black,line width=0.3mm, fit=(l) (r), inner sep=4.7pt] (r2) {};
\end{scope}

\begin{scope}[yshift=-1.85in]
\node (t)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black] {}; 
\node (b)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below=0.8 of t] {}; 
\node(ml)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below left =0.35 and 0.3 of t]{}; 
\node(mr)[circle,draw,inner sep=2pt,outer sep=1pt,line width=0.2mm,fill=black,below right=0.35 and 0.3 of t]{}; 

\path[stealth-stealth,draw,line width=0.3mm] (t)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (t)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (ml)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(mr);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(ml);
\path[stealth-stealth,draw,line width=0.3mm] (b)--(t);

\node[circle, draw=black,line width=0.3mm, fit=(t) (ml) (mr) (b), inner sep=-1.8pt] (r3) {};
\end{scope}

\path[-stealth,draw,line width=0.3mm] (r2) -- (r1);
\path[-stealth,draw,line width=0.3mm] (r3) -- (r2);
\end{tikzpicture}
}
\caption{This is a diagram of a totally comparable RPF that is not mutually possible. The mutually possible components form a total order, with the \textit{anchored component} on top.}
\label{fig:totally_comparable_rpf}
\end{figure}

\section{From Outcomes to Events}

Our next task is to upgrade \(P\) to operate on the event level. This is more difficult than it seems. For example, we may wish to declare that the probability of event \(e_1\) with respect to \(e_2\) is going to be additive on \(e_1\) as follows:
\begin{equation}
\label{eq:incorrect_additive_event_def}
P(e_1, e_2) = \sum_{h_1 \in e_1}P(h_1, e_2)
\end{equation}

Equation \ref{eq:incorrect_additive_event_def} looks uncontroversial, but it actually contradicts the fundamental axioms! If we let \(e_1 = \varnothing\), then we have an empty sum on the right hand side of the equation, and we get \(P(\varnothing, e_2) = 0\). Likewise, if we allow \(e_2\) to be empty, we get \(P(e_1, \varnothing) = P(\varnothing, e_1)^{-1}=0^{-1}=\infty\). Both of these statements make sense until you realize that \(P(\varnothing, \varnothing) = 0 = \infty\), and what's worse is that they are also equal 1 under the identity axiom!

Another problem arises when an event is \textit{internally non-comparable}, meaning that it contains outcomes \(h_1\) and \(h_2\) where \(P(h_1, h_2) = \ast\). Perhaps there are interesting things we can say about such an events, but here we will constrain ourselves to totally comparable RPFs in order to avoid such questions.

\begin{definition}
Let \(P\) be a totally comparable RPF. \(P\) can also measure the probability of two events relative to each other using the following rules:

\begin{enumerate}[(i)]
  \item \label{event_def_1} \(P(e_1, e_2)\) obeys the fundamental axioms of relative probability.
  \item \label{event_def_2} \(P(e_1, e_2)\) sums over any reference outcome \(r\), so long as the result isn't indeterminate.
    \begin{equation}
      \label{eq:event_def_ratio_match}
      P(e_1, e_2) :\cong \frac{\sum_{h_1 \in e_1} P(h_1, r)}{\sum_{h_2 \in e_2} P(h_2, r)}
    \end{equation}
\end{enumerate}
\end{definition}

Because we no longer have access to absolute probability, the best we can do is measure it relative to a \textit{reference outcome} \(r\). This ratio might be indeterminate, so we use the matching relation instead of equality. Fortunately, we can show that there exists at least one reference outcome that will constrain \(P(e_1, e_2)\) in statement \ref{eq:event_def_ratio_match} if they are non-empty.

\begin{proof} Lemma \ref{lemma:totally_comp_anchored} states that all totally comparable RPFs have anchor outcomes, and therefore (by the same argument) every event must contain outcomes that are anchors internally for that event. Choose an internal anchor \(a\) from one of the events, say \(e_1\). Then the sum \(\sum_{h_1 \in e_1} P(h_1, a)\) will be non-infinite by definition of anchors, and non-zero because \(P(a, a) = 1\) is a term in the sum. Therefore, the constraint as a whole cannot be indeterminate.

If both events are empty, then we are unable to create an anchor element, but by the identity axiom \(P(\varnothing, \varnothing) = 1\).
\end{proof}

These requirements again seem reasonable, but how can we know for sure that they provide a complete and consistent definition of \(P: \mathcal{F} \times \mathcal{F} \rightarrow \mathbb{M}^*\)? The following must be shown:

\begin{enumerate}[(i)]
  \item \label{event_def_proof_1} If two distinct values for \(r\) in statement \ref{eq:event_def_ratio_match} yield constraints on \(P\), then they must be equal.
  \item \label{event_def_proof_2} The constraint in statement \ref{eq:event_def_ratio_match} does not violate the fundamental axioms.
\end{enumerate}

\begin{proof}
For \ref{event_def_proof_1}:

Let \(r_1\) and \(r_2\) be distinct reference outcomes, and both constrain \(P(e_1, e_2)\). Then we want to check that

\begin{equation}
\label{eq:relative_event_unique}
\frac{\sum_{h_1 \in e_1} P(h_1, r_1)}{\sum_{h_2 \in e_2} P(h_2, r_1)} = \frac{\sum_{h_1 \in e_1} P(h_1, r_2)}{\sum_{h_2 \in e_2} P(h_2, r_2)}
\end{equation}

Neither expression is a wildcard, and none of the individual terms are either. The key to this argument is in looking at the value of \(P(r_1, r_2)\).

Assume \(P(r_1, r_2) = 0\). 

If \(\sum_{h_1 \in e_1} P(h_1, r_1)\) is not infinite, then \(\sum_{h_1 \in e_1} P(h_1, r_2)\) must be zero. The same argument applies to \(\sum_{h_2 \in e_2} P(h_2, r_2)\). Since they can't both be zero, we can say that one of the sums on the left hand side is infinite, so that \(P(e_1, e_2)\) is either \(\infty\) or 0. Let's say it is 0. Then \(\sum_{h_1 \in e_1} P(h_1, r_1) = 0\) and \(\sum_{h_2 \in e_2} P(h_2, r_1) = \infty\) and by the argument above \(\sum_{h_1 \in e_1} P(h_1, r_2) = 0\). Because the right hand side is not \(\ast\) - it must resolve to zero as well. The same agument holds for \(P(e_1, e_2) = \infty\).

By an analogous argument, equation \ref{eq:relative_event_unique} must also hold when \(P(r_1, r_2) = \infty\).

So now we can assume that \(P(r_1, r_2) \notin \{0, \infty\} \). Multiply the left hand side of equation \ref{eq:relative_event_unique} by \(1 = \frac{P(r_1, r_2)}{P(r_1, r_2)}\) and distribute to get:

\[\frac{\sum_{h_1 \in e_1} P(h_1, r_1) \cdot P(r_1, r_2)}{\sum_{h_2 \in e_2} P(h_2, r_1) \cdot P(r_1, r_2)} = \frac{\sum_{h_1 \in e_1} P(h_1, h_2^*)}{\sum_{h_2 \in e_2} P(h_2, h_2^*)}\]

For \ref{event_def_proof_2}:

The identity, inverse, and composition axioms follow from the fact that statement \ref{eq:event_def_ratio_match} is a ratio with identical expressions for \(e_1\) in the numerator and \(e_2\) in the denominator. Therefore, if it resolves it is just a ratio of positive numbers - which can be shown to follow the 3 axioms.
\end{proof}

\begin{theorem}
If events \(e_1\) and \(e_2\) are not both empty, the following formula for calculating the relative probability of events is true:
\[P(e_1, e_2) = \sum_{h_1 \in e_1} \frac{1}{\sum_{h_2 \in e_2} P(h_2, h_1)}.\]
\end{theorem}

\begin{proof}
Find a suitable reference outcome \(h\) and multiply by \(1 = \frac{P(h_1, r)}{P(h_1, r)}\).
\[\sum_{h_1 \in e_1} \frac{1}{\sum_{h_2 \in e_2} p(h_2, h_1)} :\cong \sum_{h_1 \in e_1} \frac{P(h_1, r)}{\sum_{h_2 \in e_2} P(h_2, h_1) P(h_1, r)} = \frac{\sum_{h_1 \in e_1} P(h_1, r)}{\sum_{h_2 \in e_2} P(h_2, r)}\]

Since both \(P(e_1, e_2)\) and the formula above match the same thing which is not \(\ast\) for appropriate reference r, they must be equal.
\end{proof}

We then derive the absolute probability function as
\[P(e) = P(e, \Omega) = \sum_{h \in e} \frac{1}{\sum_{h' \in \Omega}p(h', h)}\]

\begin{theorem}
\label{thm:empty_event_impossible}
The empty event \(\varnothing\) has probability 0 relative to any non-empty event.
\end{theorem}

\begin{proof}
Let \(e\) be a non-empty event, and let \(h\) be an outcome in \(e\).

\[P(\varnothing, e) :\cong \frac{\sum_{h_1 \in \varnothing} P(h_1, h)}{\sum_{h_2 \in e} P(h_2, h)} = \frac{0}{\sum_{h_2 \in e} P(h_2, h)}\]

The sum \(\sum_{h_2 \in e} P(h_2, h)\) cannot itself be zero because \(P(h, h)\) is one of its terms. Therefore, \(P(\varnothing, e) = 0\)
\end{proof}

\section{Composing Relative Probability Functions}

Let \(P_0, P_1, ..., P_{K-1}\) be relative probability functions. Each of these probability functions have a unique outcome space. Let \(P_k\) measure relative probability on outcome space \(\Omega_k\), so that \(P_k: \Omega_k \times \Omega_k \rightarrow \mathbb{M}^{\ast}\).

We can combine all of these relative probability functions together with a top level probability function \(P_\top\)\footnote{Pronounced \quotes{P-Top}.} with outcome space \(\Omega_\top = \{\Omega_0, \Omega_1, ... \Omega_{K- 1}\}\). The outcome space is heirarchical as shown in figure \ref{fig:compose_rpfs}.

\begin{figure}[h]
\begin{tikzpicture}
\node {\(\Omega_\top\)} [sibling distance = 4cm]
  child {node {\(\Omega_0\)}  [sibling distance = 1cm]
    child {node {\(h_{0, 0}\)}}
    child {node {\(h_{0, 1}\)}}
    child {node {\dots}}
    child {node {\(h_{0, |\Omega_0| - 1}\)}}
  }
  child {node {\(\Omega_1\)}  [sibling distance = 1cm]
    child {node {\(h_{1, 0}\)}}
    child {node {\dots}}  
  }
  child {node {\dots}  [sibling distance = 1cm]
    child {node {\dots}}
    child {node {\dots}}
  }
  child {node {\(\Omega_{K-1}\)}   [sibling distance = 2cm]
    child {node {\dots}}
    child {node {\(h_{K-1, |\Omega_{K-1}| - 1}\)}}
  };
\end{tikzpicture}
\caption{A tree diagram for a set of RPFs being composed by a top-level RPF.}
\label{fig:compose_rpfs}
\end{figure}

Now let \(\Omega\) be the set of all outcomes \(\Omega_0 \cup \Omega_1 \cup \dots \Omega_{K-1}\). We can create a new RPF - just called \(P\) acting on \(\Omega\) - with the following rules:

\begin{itemize}
\item If the two outcomes fall under the same component, then their relative probabilities do not change:

\begin{equation}
\label{rpf_composition_same_branch}
P(h_{k, i}, h_{k, j}) = P_k(h_{k, i}, h_{k, j})
\end{equation}

\item If the two outcomes fall under different components, then their relative probabilities are given as follows.

\begin{equation}
\label{eq:rpf_composition_different_branch}
P(h_{k_1, i}, h_{k_2, j}) = P_{k_1}(h_{k_1, i}, \Omega_{k_1}) \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_2}) \cdot P_{k_2}(\Omega_{k_2}, h_{k_2, j})
\end{equation}
\end{itemize}

Note the use of the composition property to traverse up and down the tree. One could of course imagine this tree being many levels, and having a different height for each branch.

\begin{theorem}
\(P\) respects the fundamental axioms.
\end{theorem}

\begin{proof}
Identity is obvious because an outcome is on the same component as itself, so we can use equation \ref{rpf_composition_same_branch} to get \(P(h_{k, i}, h_{k, i}) = P_k(h_{k, i}, h_{k, i}) = 1\)

The inverse and composition laws must be true if both inputs are in the same component, because that component already follows the axioms. We now look at two inputs are from different components.

The inverse law can be proven by calculation.
\begin{equation}
\begin{aligned}
P(h_{k_1, i}, h_{k_2, j})^{-1} &= (P_{k_1}(h_{k_1, i}, \Omega_{k_1}) \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_2}) \cdot P_{k_2}(\Omega_{k_2}, h_{k_2, j}))^{-1} \\
& = P_{k_1}(h_{k_1, i}, \Omega_{k_1})^{-1} \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_2})^{-1} \cdot P_{k_2}(\Omega_{k_2}, h_{k_2, j})^{-1} \\
& = P_{k_1}(\Omega_{k_1}, h_{k_1, i}) \cdot  P_{\top}(\Omega_{k_2}, \Omega_{k_1}) \cdot P_{k_2}(h_{k_2, j}, \Omega_{k_2}) \\
& = P_{k_2}(h_{k_2, j}, \Omega_{k_2})\cdot  P_{\top}(\Omega_{k_2}, \Omega_{k_1}) \cdot P_{k_1}(\Omega_{k_1}, h_{k_1, i}) \\
& = P(h_{k_2, j}, h_{k_1, i})
\end{aligned}
\end{equation}

Composition can be shown similarly - now naming the 3 separate indecies in components \(k_1, k_2, k_3\) as \(i_1, i_2, i_3\) respectively.
\begin{equation}
\begin{aligned}
& P(h_{k_1, i_1}, h_{k_2, i_2}) \cdot P(h_{k_2, i_2}, h_{k_3, i_3}) \\
& :\cong P_{k_1}(h_{k_1, i_1}, \Omega_{k_1}) \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_2}) \cdot \textcolor{red}{P_{k_2}(\Omega_{k_2}, h_{k_2, i_2}) \cdot  P_{k_2}(h_{k_2, i_2}, \Omega_{k_2})} \cdot  P_{\top}(\Omega_{k_2}, \Omega_{k_3}) \cdot P_{k_3}(\Omega_{k_3}, h_{k_3, i_3}) \\
& :\cong P_{k_1}(h_{k_1, i_1}, \Omega_{k_1}) \cdot  \textcolor{red}{P_{\top}(\Omega_{k_1}, \Omega_{k_2}) \cdot  P_{\top}(\Omega_{k_2}, \Omega_{k_3})} \cdot P_{k_3}(\Omega_{k_3}, h_{k_3, i_3}) \\
& :\cong P_{k_1}(h_{k_1, i_1}, \Omega_{k_1}) \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_3}) \cdot P_{k_3}(\Omega_{k_3}, h_{k_3, i_3}) \\
& :\cong P_{k_1}(h_{k_1, i_1}, h_{k_3, i_3})
\end{aligned}
\end{equation}
\end{proof}

\begin{theorem}
\(P\) is totally comparable if and only if the following are true:

\begin{enumerate}
\item \(P_{\top}\) is totally comparable.
\item For all \(k \in \{0, 1, ..., K - 1\}\), \(P_k\) is totally comparable.
\item All components except at most one are totally mutually possible.
\item If there is a component that is not totally mutually possible, then every element of \(P_{\top}\) possible with respect to that component.
\end{enumerate}
\end{theorem}

\begin{proof}
If all the components are totally comparable, then any two outcomes in the same component are always going to be comparable in the overall RPF. We only need to prove that outcomes in \textbf{different} components are comparable. Starting with equation \ref{eq:rpf_composition_different_branch},

\begin{equation}
P(h_{k_1, i}, h_{k_2, j}) = P_{k_1}(h_{k_1, i}, \Omega_{k_1}) \cdot  P_{\top}(\Omega_{k_1}, \Omega_{k_2}) \cdot P_{k_2}(\Omega_{k_2}, h_{k_2, j})
\end{equation}

The only way that we can get \(P(h_{k_1, i}, h_{k_2, j}) = \ast\) is if there are both \(0\) and \(\infty\) as factors on the right hand side.

Because there is at most one component with outcomes impossible with respect to that component, we can say that either \(P_{k_1}(h_{k_1, i}, \Omega_{k_1}) = 0\) or \(P_{k_2}(h_{k_2, j}, \Omega_{k_2}) = 0\), or possibly neither, but not both.

Neither can be infinite either by the definition of the event level in equation \ref{eq:event_def_ratio_match}. Here we look at the factor \(P_{k_1}(h_{k_1, i})\) and use \(k_1\) itself as the reference outcome.
\[
P_{k_1}(h_{k_1, i}, \Omega_{k_1}) :\cong \frac{\sum_{h_1 \in \{k_1\}} P(h_1, k_1)}{\sum_{h \in \Omega_{k_1}} P(h_2, k_1)} = \frac{1}{\sum_{h \in \Omega_{k_1}} P(h_2, k_1)}
\]

The sum in the denominator cannot be zero since \(P(k_1, k_1) = 1\) will be one of its terms.

If the term \(P_{k_1}(h_{k_1, i}) = 0\), then the only way the entire right hand side can be \(\ast\) is if \(P_{\top}(\Omega_{k_1}, \Omega_{k_2}) = \infty\). But this can't be true because we assumed that \(\Omega_{k_2}\) is possible with respect to \(\Omega_{k_1}\), the sole component with impossible outcomes!

An analogous argument can be made if \(P_{k_2}(h_{k_2, j}, \Omega_{k_2}) = 0\).

Therefore, the right hand side of the equation is not \(\ast\) and \(P\) is totally comparable.

In the opposite direction, we can show that if any of the conditions are broken, then \(P\) is not totally comparable. Breaking any of the first two conditions would introduce an explicit \(\ast\) into equation \ref{eq:rpf_composition_different_branch}. If there are multiple components with impossible outcomes, then it would introduce a \(0\) into the first term of equation \ref{eq:rpf_composition_different_branch} and an \(\infty\) into the third term, yielding \(\ast\).

And finally, if only the fourth condition is broken, it would introduct a 0 into the first term of equation \ref{eq:rpf_composition_different_branch} and an \(\infty\) into the \textbf{second} term of equation \ref{eq:rpf_composition_different_branch}.

Therefore, if any of these conditions are broken, \(P\) is \textbf{not} totally comparable.
\end{proof}

\section{Topology and Limits in Relative Probability Space}
\label{section:topology}

Mathematics can be used to model the real world even through seemingly impossible ideas. For example, we might believe that a certain natural process cannot repeat an infinite number of times - that it is just not something allowed by the physical limitations of our universe. And even so, we might still speak of \quotes{infinite iterations} in order to get a bound or estimate on what that system will look like in \quotes{the long run}. One of the benefits of relative probability spaces is their properties with respect to limits. To this end, we prove here that when we take limits of totally comparable RPFs, the result will also be totally comparable.

This effort caps off a significant argument in favor of totally comparable RPFs. They hold their information under the operation of limits, while absolute probability does not.

There is some background in topology\footnote{See Mendelson (1990) \cite{mendelson} and Bradley et al. (2020) \cite{bradley} for texts with formal definitions and theorems.} required for this section.

\subsection{RPF Space and Compactness}

Because the set of absolute distributions is embedded in \(\mathbb{R}^K\), its topological properties are well understood. The simplex is closed, bounded, and compact. Practically, this means that any sequence of points on the simplex will converge to one or more points on the simplex allowing both pure and applied practicioners to talk about limit and boundary conditions.

This strategy fails for relative probabilities, because there is no obvious way to embed an RPF into euclidean space\footnote{Though it may be possible! See section \ref{section:euclidean_embedding}}. The relative probability space is more complicated, because at the corners and edges of the simplex lurk entire subspaces where zero-probability outcomes are still being compared in different ways.

\begin{definition}
\(\text{RPF}^{\ast}(K)\) is the set of relative probability functions of size K (where \(\Omega = \{0, 1, ..., K - 1\}\)). Likewise \(\text{RPF}(K)\) is the set of all totally comparable RPFs of size K.
\end{definition}

If the \(\text{RPF}(K)\) is compact, then information about the relative probabilites of events are preserved even as they approach zero relative to another event.

In order to prove compactness, we first must define a \textit{topology} on \(\text{RPF}(K)\). This starts with finding a \textit{basis of open sets}.

The notion of an open set changes when a topological space is restricted to a lower dimention. For example, on the real number line \(\mathbb{R}\), we take the open interval (0, 1) as an open set. However, once this is embedded into \(\mathbb{R}^2\), it is now a line segment in a plane and no longer open (see figure \ref{fig:sub_topology}). It can be thought of as the restriction of an open set on \(\mathbb{R}^2\) to \(\mathbb{R}\). For example, the set \(\{(x, y): x \in (0, 1)\;  \text{and}\;  y \in (-\epsilon, +\epsilon)\}\) given an \(\epsilon > 0\) is such an open set on \(\mathbb{R}^2\).

\begin{figure}[h]
\centering
\resizebox{0.4\textwidth}{!}{
\begin{tikzpicture} %episilon diagram
\draw[<->] (-0.7,0)--(5,0) node[right] {\(x\)};
\draw[<->] (0,-1,0)--(0,1) node[above] {\(y\)};
\draw[densely dashed,red,thick] (0,-0.5) rectangle (4.5,0.5);
\draw[draw=blue,thick] (0,0) circle (2pt) node[above left=-0.1,font=\ttfamily\tiny] {0};
\draw[draw=blue,thick] (4.5,0) circle (2pt) node [above right=-0.1,font=\ttfamily\tiny] {1};

\node[left=-0.05,font=\ttfamily\footnotesize] at (0,-0.5) {\(-\varepsilon\)};
\node[left=-0.05,font=\ttfamily\footnotesize] at (0,0.5) {\(\varepsilon\)};
\node[circle,inner sep=1pt] at (0,0) (0){};
\node[circle,inner sep=1pt] at (4.5,0) (1){};
\path[draw,blue,thick] (0)--(1);
\end{tikzpicture}
}
\caption{The small box that is the interior of the dotted rectangle is an open set in \(\mathbb{R}^2\), and therefore its restriction to \(\mathbb{R}\) - the line segment - is an open set in \(\mathbb{R}\). But the line segment is not open in \(\mathbb{R}^2\).}
\label{fig:sub_topology}
\end{figure}


Likewise, an open set on a relative probability space restricted on several outcomes might not be an open set on the relative probability spaces for all of \(\Omega\).

We start by looking at RPFs with \(K = 2\). Fortunately, we find a totally comparable RPF that corresponds 1:1 with the magnitude space.

\begin{theorem}
Let \(\Omega = \{h_1, h_2\}\) have two elements, with relative probability function \(P\). Then, \(P\) is completely determined by \(P(h_1, h_2)\).
\end{theorem}

\begin{proof}
Let \(q = P(h_1, h_2)\). By the inverse symmetric property, \(P(h_2, h_1) = q^{-1}\). These values completely determine \(P\) on the outcome level.
\end{proof}

This gives us both a topology and a compactness proof for K = 2 for free because \(\text{RPF}(2)\) is isomorphic to \(\mathbb{M}\) which already has a natural topology. Its basis for open sets are the open intervals of \(\mathbb{B}\), including those intervals that include 0 and \(\infty\). For \(K > 2\), we will need more powerful tools.

\subsection{Open Patches}

We now develop a notion of open patches, which will be a basis of open sets on the space \(\text{RPF}(K)\).

\begin{definition}
An \textit{interior open patch} of \(\text{RPF}(K)\) is one of the following:

\begin{enumerate}
  \item If \(K = 2\), a subset parameterized by an interior open interval of magnitudes. \(\{P | a < P(h_1, h_2) < b\}\) for some \(a, b \in \mathbb{M}\) 
  \item If \(K > 2\), a composition of interor patches with composing function \(P_{\top}\) also being an interior patch.
\end{enumerate}
\end{definition}

Interior open patches contain only totally mutually possible functions as illustrated in figure \ref{fig:interior_open_patch}.

\begin{figure}[h]
\centering
\resizebox{0.4\textwidth}{!}{
\begin{tikzpicture}
\node {\(P_\top\)} [sibling distance = 3cm]
  child {node {\(P_0\)}  [sibling distance = 3cm]
    child {node {A}}
    child {node {B}}
  }
  child {node  {C}};
\end{tikzpicture}
}
 \hspace{3em}
\resizebox{0.4\textwidth}{!}{
\begin{tikzpicture} %Triangle 3
\draw[thick, dashed,red] (0,{2*sqrt(3)})--(1,0) node [font=\tiny\bf] {)};
\draw[thick, dashed,red] (0,{2*sqrt(3)})--(0,0) node [font=\tiny\bf] {(};
\path[draw=none,pattern=crosshatch, pattern color=darkgreen] (0,1.2)--(0.66,1.2)--(0.43,2)--(0,2)--cycle;

\node[rectangle,rotate=45,draw=none,pattern=crosshatch, pattern color=darkgreen] at (1.2,3) (P){};

\node(Q)[rectangle,draw=none,align=center,below right=-0.3 and 0.1 of P,font=\ttfamily\footnotesize] {Interior Open\\ Patch};

\draw[thick] (-2,0)--(2,0)--(0,{2*sqrt(3)})--cycle;

\node (A) at (-2,0) [below] {\(A\)}; 
\node (B) at (2,0) [below] {\(B\)}; 
\node (C) at (0,{2*sqrt(3)}) [above] {\(C\)}; 

\draw[thick, dashed,red] (0.8,2)--(-3,2) node[font=\tiny\bf,rotate=90] {)};
\draw[thick, dashed,red] (1.2,1.2)--(-3,1.2) node[font=\tiny\bf,rotate=90] {(};

\draw[thick, dashed,red] (1,-0.1)--(1,-0.9);
\draw[thick, dashed,red] (0,-0.1)--(0,-0.9);

\draw[thick,|-|] (-3,0)--(-3,{2*sqrt(3)}) node[right,pos=0.9] {\(P_{T}\)};

\draw[thick,|-|] (-2,-1)--(2,-1) node[above,pos=0.9] {\(P_{0}\)} node[red,font=\tiny\bf] at (1,-1) {)}
node[red,font=\tiny\bf] at (0,-1) {(};
\end{tikzpicture}
}
\caption{An interior open patch captures a contiguous set inside the probability simplex. Above is an example of a composite probability distribution where the diagram on the left shows how \(P_\top\) and \(P_0\) are composed, and the right shows how the interior segments on both interact to form a patch. }
\label{fig:interior_open_patch}
\end{figure}

\begin{definition}
A \textit{facet\footnote{A facet of a simplex is a subset where one parameter is equal to zero - equivalent to a face on a 3D object.} patch} of \(\text{RPF}(K)\) is one of the following:

\begin{enumerate}
  \item If \(K = 2\), an interval of the form \(\{P | 0 < P(h_1, h_2) < a\}\) for some \(a \in \mathbb{M}\) 
  \item If \(K > 2\), a composition where \(P_{\top}\) is drawn from an interior open patch, and all but one of the components are drawn from interior open patches. The final component - the \textit{facet component} - is itself drawn from a facet patch.
\end{enumerate}
\end{definition}

\begin{definition}
An \textit{exterior open patch} is a one of the following:

\begin{enumerate}
  \item A facet patch.
  \item A composition where \(P_{\top}\) is a facet patch. The \textit{facet component} is itself drawn from any open patch, and all the other components are drawn from interior open patches.
\end{enumerate}
\end{definition}

As seen in figure \ref{fig:exterior_open_patch}, exterior open patches touch the hyperfaces (facets) of the simplex as well as the vertices and edges.  As the number of dimentions increases and the composition diagram changes, more permutations are possible.

\begin{figure}[h]
\centering
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture} %Triangle 2
\draw[thick, dashed,red] (0,{2*sqrt(3)})--(-1,0) node [font=\tiny\bf] {)};
\path[draw=none,pattern=crosshatch, pattern color=darkgreen] (-1.3,1.2)--(-0.66,1.2)--(-0.43,2)--(-0.86,2)--cycle;

\draw[thick] (-2,0)--(2,0)--(0,{2*sqrt(3)})--cycle;

\node (A) at (-2,0) [below] {\(A\)}; 
\node (B) at (2,0) [below] {\(B\)}; 
\node (C) at (0,{2*sqrt(3)}) [above] {\(C\)}; 

\draw[thick, dashed,red] (0.8,2)--(-3,2) node[font=\tiny\bf,rotate=90] {)};
\draw[thick, dashed,red] (1.2,1.2)--(-3,1.2) node[font=\tiny\bf,rotate=90] {(};

\draw[thick, dashed,red] (-1,-0.1)--(-1,-0.9);
\draw[thick, darkgreen] (-1.3,1.2)--(-0.85,2);

\draw[draw=darkgreen] (-1.3,1.2) circle (2pt);
\draw[draw=darkgreen] (-0.85,2) circle (2pt);

\draw[thick,|-|] (-3,0)--(-3,{2*sqrt(3)}) node[right,pos=0.9] {\(P_{T}\)};

\draw[thick,|-|] (-2,-1)--(2,-1) node[above,pos=0.9] {\(P_{0}\)} node[red,font=\tiny\bf] at (-1,-1) {)};
\end{tikzpicture}
}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture} %Triangle 1
\draw[thick, dashed,red] (0,{2*sqrt(3)})--(-1,0) node [font=\tiny\bf] {)};
\path[draw=none,pattern=crosshatch, pattern color=darkgreen] (-0.36,2.2)--(0,{2*sqrt(3)})--(-0.75,2.2)--cycle;

\draw[thick] (-2,0)--(2,0)--(0,{2*sqrt(3)})--cycle;

\node (A) at (-2,0) [below] {\(A\)}; 
\node (B) at (2,0) [below] {\(B\)}; 
\node (C) at (0,{2*sqrt(3)}) [above] {\(C\)}; 
\draw[thick, dashed,red] (0.8,2.2)--(-3,2.2) node[font=\tiny\bf,rotate=90] {(};
\draw[thick, dashed,red] (-1,-0.1)--(-1,-0.9);
\draw[thick, darkgreen] (-0.73,2.2)--(0,{2*sqrt(3)});

\draw[draw=darkgreen] (-0.73,2.2) circle (2pt);

\draw[draw=darkgreen] (0,{2*sqrt(3)}) circle (2pt);

\fill[darkgreen] (0,{2*sqrt(3)}) + (0, 2pt) arc (90:270:2pt);

\draw[thick,|-|] (-3,0)--(-3,{2*sqrt(3)}) node[right,pos=0.9] {\(P_{T}\)};

\draw[thick,|-|] (-2,-1)--(2,-1) node[above,pos=0.9] {\(P_{0}\)} node[red,font=\tiny\bf] at (-1,-1) {)};
\end{tikzpicture}
}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture} %Triangle 1
\path[draw=none,pattern=crosshatch, pattern color=darkgreen] (0.36,2.2)--(0,{2*sqrt(3)})--(0,2.2)--cycle;

\draw[thick] (-2,0)--(2,0)--(0,{2*sqrt(3)})--cycle;

\node (A) at (-2,0) [below] {\(A\)}; 
\node (B) at (2,0) [below] {\(B\)}; 
\node (C) at (0,{2*sqrt(3)}) [above] {\(C\)}; 
\draw[thick, dashed,red] (0.8,2.2)--(-3,2.2) node[font=\tiny\bf,rotate=90] {(};

\draw[draw=darkgreen] (0,{2*sqrt(3)}) circle (2pt);

\fill[darkgreen] (0,{2*sqrt(3)}) + (0, 2pt) arc (90:270:2pt);

\draw[thick, dashed,red] (0,{2*sqrt(3)})--(1,0) node [font=\tiny\bf] {)};
\draw[thick, dashed,red] (0,{2*sqrt(3)})--(0,0) node [font=\tiny\bf] {(};

\draw[thick,|-|] (-3,0)--(-3,{2*sqrt(3)}) node[right,pos=0.9] {\(P_{T}\)};
\draw[thick,|-|] (-2,-1)--(2,-1) node[above,pos=0.9] {\(P_{0}\)};

\draw[thick, dashed,red] (1,-0.1)--(1,-0.9);
\draw[thick, dashed,red] (0,-0.1)--(0,-0.9);

\draw[thick,|-|] (-2,-1)--(2,-1) node[above,pos=0.9] {\(P_{0}\)}
node[red,font=\tiny\bf] at (1,-1) {)}
node[red,font=\tiny\bf] at (0,-1) {(};
\end{tikzpicture}
}
\caption{Exterior open patches. On the left is the facet patch, because it only touches a side (facet) of the simplex and not a corner. In the center is an exterior open patch where the facet component \(P_0\) is itself a facet patch (touching an edge and a corner), and on the right is an exterior open patch that touches a corner only because \(P_0\) is an interior open patch. Note that the point containing the corner at \(C\) in the middle and third diagram is only half filled because the patch contains some values where \(P(C) = 1\) and not others, depending on the relative probability between \(A\) and \(B\).}
\label{fig:exterior_open_patch}
\end{figure}

\begin{definition}
An \textit{open patch} is a subset of \(\text{RPF}(K)\) that is either an interior or exterior open patch.
\end{definition}

Now let the open patches be the bases for an open set thus defining a topology on \(\text{RPF}(K)\).

\begin{definition}
An \textit{open set} of \(\text{RPF}(K)\) is any (potentially infinite) union of open patches on \(\text{RPF}(K)\), or any finite intersection of open patches on \(\text{RPF}(K)\).
\end{definition}

\subsection{Compactness}

The compactness proof for \(\text{RPF}(K)\) will be sketched here for brevity. First, we need a few lemmas.

\begin{lemma}
\label{lem:withhold_possible_outcome}
Let \(h\) be an outcome, and let q be a number such that \(0 < q \leq 1\). The region of \(\text{RPF}(K)\) where the \(P(h) = q\) is isomorphic to \(\text{RPF}(K-1)\).
\end{lemma}

\begin{proof}
If \(P(h) > 0\) then it is in the anchored equivalence class of mutually possibility. If \(P(h) = 1\) then the outcome \(h\) can be appended above any function in \(\text{RPF}(K - 1)\), and if \(P(h) < 1\) then \(h\) can be appended into the anchored equivalence class of any function in \(\text{RPF}(K-1)\). In both cases, a separate \(h\) with a given absolute probability can be appended to anything in \(\text{RPF}(K-1)\) to produce an element of \(\text{RPF}(K)\), with all elements of \(\text{RPF}(K)\) accounted for.
\end{proof}

\begin{lemma}
\label{lem:change_one_anchor_outcome}
Let \(h\) be an outcome, and let \(P(h) > 0\). Given a number \(0 < q \leq 1\), there is a unique RPF \(P'\) which is equal to \(P\) when evaluated on outcomes \(\neq h\) and such that \(P'(h) = q\).
\end{lemma}

\begin{proof}
This will be a proof by construction. Let \(P'\) be the new distribution where \(h\) remains an anchor outcome, but its absolute probability has been changed to \(q\). For any anchor outcome \(a\), we set \(P'(a, h) = P(a, h) \cdot q \cdot P(a)^{-1} \)

For any non-anchor outcome \(b\), we note that \(P(b, h) = 0\) and therefore also \(P'(b, h) = 0\), otherwise \(b\) would now be possible relative to other anchors where it wasn't before.
\end{proof}

\begin{lemma}
\label{lem:open_anchor_outcome}
Let \(h\) be an outcome, and let \(P(h) > 0\). Any open patch of \(\text{RPF}(K)\) that contains \(P\), also contains, for some open interval on \(\mathbb{M}\) containing \(P(h)\), all the values \(P'\) constructed through lemma \ref{lem:change_one_anchor_outcome} by setting \(P'(h) = q\) for any \(q\) in that open interval.
\end{lemma}

\begin{proof}
TODO
\end{proof}

\begin{theorem}
\(\text{RPF}(K)\) is \textit{compact}, meaning that for every open cover of it, there is a finite subcover.
\end{theorem}

We will provide a sketch for the compactness proof here.

\begin{proof}
This is an inductive proof where we assume that the theorem is true for all \(k < K\) and then prove that it is true for \(K\).

If \(K \in {0, 1}\) then \(\text{RPF}(K)\) is finite and singular (either the empty RPF or unit RPF respectively). These are obviously compact. If \(K = 2\) then we have the topology of \(\mathbb{M}\) which is also compact (thanks to the \(\infty\) element).

Now we assume that \(K > 2\).

Consider the region of \(\text{RPF}(K)\) where a specific outcome is required to be the largest (or possibly ties for largest). In other words, for this special outcome \(h\), look at the region of \(\text{RPF}(K)\) where \(P(h', h) \leq  1\) for any other outcome \(h'\). There is one region for every outcome - \(K\)such regions overall - and collectively they cover \(\text{RPF}(K)\) but they are not disjoint. In fact, the uniform distribution belongs to all \(K\) regions!

If we can show that an open cover on \(\text{RPF}(K)\) has a finite subcover on a region, then it must have a finite subcover overall because there are a finite number of regions.

So consider one such region, and let \(h\) be the largest outcome from that region. \(h\) is an anchor element, which means that \(P(h, h') > 0\) for all \(h'\) when in fact it is \(\geq 1\). If \(P(h) = q\), then we know that \(\frac{1}{k} \leq q \leq 1\). By lemma \ref{lem:withhold_possible_outcome}, the rest of the distribution is isomorphic to \(\text{RPF}(K - 1)\) - and therefore compact by our inductive assumption. Therefore, there is a finite subcover covering all elements where \(P(h) = q\).

By lemma \ref{lem:open_anchor_outcome}, finite subcover will not just include cases where \(P(h) = q\), but will hold for some open interval around q as well.

So for each q where \(\frac{1}{K} \leq q \leq 1\), we have a finite subcover, with each subcover covering some open set around q. Because \([\frac{1}{K}, 1]\) is compact, all of these open sets around q have a finite subcover for the whole segment. This means that only finitely many of these q-based subcovers are required, and with each one being finite we have finitely many open sets for the entire region where \(h\) is the largest outcome.
\end{proof}

\section{Future Work}
\subsection{Expansions to infinite spaces}
The obvious extention to this work is to expand relative probability to a generalized space which may be infinite, and thus capture all of the variety of probability distributions that one might wish to study and apply. This would start by modifying statement \ref{eq:event_def_ratio_match} to ask for an additive property. The relative probability function would then become a \textit{relative probability distribution}.

This raises certain question which - while decisions have been made in prior work and certainly in measure theory - should be open to discussion.
\begin{enumerate}
\item Do we need to keep countable additivity as set forth in the Kolmogorov axioms, or can we relax this to allow for a fair countable lottery? RPFs would provide a great way to analyze the fair countable lottery and this should be exploited!
\item If we derive a notion of probability density, then can these densities at a particular pair of events be used to compare the relative probability of those events? What specific properties of the relative probability distribution are required to make this work?
\end{enumerate}

It appears possible to use these ideas to create a unified version of the Hausdorff measure - which finds the size of an object given its dimention. Instead of considering it to be multiple measures - we can have a single measure where bounded sets of equal dimention are mutually possible, and smaller-dimentional objects are always mutually impossible with respect to a larger dimentional objects.

\subsection{Relationship to Category Theory}

Category theorists will instantly recognize that an RPF describes a category. This construction can be analyzed and approached through the lens of category theory. Specifically, and RPF describes something called a \textit{thin category} where any pair of objects have at most one morphism connecting them (per direction).

\begin{center}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture} %Commute Diagram
[node distance=0.8in]
\tikzset{
  shift left/.style ={commutative diagrams/shift left={#1}},
  shift right/.style={commutative diagrams/shift right={#1}},
  com/.style={circle,draw=none,inner sep=1pt,font=\LARGE}
}
\node (C) [com] {\(C\)};
\node (B) [com, above right = of C] {\(B\)};
\node (A) [com, above left = of C] {\(A\)};

\path[-stealth,thick,shift right=0.7ex] (C) edge node[right] {6} (A);
\path[-stealth,thick,shift right=0.7ex] (C) edge node[right] {3} (B);
\path[-stealth,thick,shift right=0.7ex] (B) edge node[above] {2} (A);


\path[-stealth,thick,shift right=0.7ex] (A) edge node[left,xshift=-0.6ex] {\(\frac{1}{6}\)} (C);
\path[-stealth,thick,shift right=0.7ex] (B) edge node[left,yshift=0.6ex] {\(\frac{1}{3}\)} (C);
\path[-stealth,thick,shift right=0.7ex] (A) edge node[below] {\(\frac{1}{2}\)} (B);

\path[<-,every loop/.style={looseness=5},thick] (A)
         edge  [in=170,out=90,loop,below] node {1} (); 
\path[<-,every loop/.style={looseness=5},thick] (B)
         edge  [in=80,out=0,loop,below] node {1} (); 
\path[<-,every loop/.style={looseness=9},thick,rotate=180] (C)
         edge  [in=-235,out=-315,loop,above] node {1} ();

%\path[-stealth,thick] (A) edge [loop above] node {1} ();
\end{tikzpicture}
}
\end{center}

The recent work of Censi et al.\cite{censi} concerns negative information in categories, which corresponds to the wildcard element \(\ast\). It represents regions of the probability function that remain uncomparable. This work could be used to subsume and develop the indeterminate wildcard concept.

\subsection{Embedding in Euclidean Space}
\label{section:euclidean_embedding}

Absolute probability functions have this advantage where they can be embedded into a simplex in \(\mathbb{R}^K\). For relative probability functions, it is not so straightforward. However, it should still be possible to embed finite RPFs into euclidean space. For example, the space \(\text{RPF}(3)\) can be mapped as a hexagon, where each point can be assigned a probability based on its distance between two parallel sides, which exist for each outcome.

\begin{center}
\resizebox{0.25\textwidth}{!}{
\begin{tikzpicture}
\node[regular polygon,draw,minimum size=4cm,line width=0.5mm,regular polygon sides = 6] (p) at (0,0) {};

\node at (p.corner 1) [anchor=360/6*(1-1)+270] {};
\node at (p.corner 2) [anchor=360/6*(2-1)+270] {};
\node at (p.corner 3) [anchor=360/6*(3-1)+270] {};
\node at (p.corner 4) [anchor=360/6*(4-1)+270] {};
\node at (p.corner 5) [anchor=360/6*(5-1)+270] {};
\node at (p.corner 6) [anchor=360/6*(6-1)+270] {};

\node[regular polygon,draw=none,rotate=30,minimum size=3.4cm,line width=0.5mm,regular polygon sides = 6] (A) at (0,0) {};
\node at (A.corner 1) [anchor=360/6*(1-1)+270] {\(P(a)=1\)};
\node at (A.corner 2) [anchor=360/6*(2-1)+230,rotate=58] {\(P(c)=0\)};
\node at (A.corner 3) [anchor=360/6*(4-1)+260,rotate=-60] {\(P(b)=1\)};
\node at (A.corner 4) [anchor=360/6*(4-1)+270] {\(P(a)=0\)};
\node at (A.corner 5) [anchor=360/6*(5-1)+230,rotate=55] {\(P(c)=1\)};
\node at (A.corner 6) [anchor=360/6*(7-1)+260,rotate=-60] {\(P(b)=0\)};

\draw[fill=black] (0.5,0.6) circle (3pt);

\draw[line width=0.5mm,dashed] (0.5,0.6) -- ($(p.corner 1)!(0.5,0.6)!(p.corner 6)$);
\draw[line width=0.5mm,dashed] (0.5,0.6) -- ($(p.corner 2)!(0.5,0.6)!(p.corner 3)$);
\draw[line width=0.5mm,dashed] (0.5,0.6) -- ($(p.corner 5)!(0.5,0.6)!(p.corner 4)$);
\end{tikzpicture}
}
\end{center}

In this case, the probability triangle has been truncated. For higher order simplices, this appears to become exceedingly unweidly unless some simplifying trick is developed. If it is successfully done, then the topological properties of \(\text{RPF}(K)\) fall into place easily.

It is also possible to hack a metric space for \(\text{RPF}(K)\) by assigning distances to elements based on their euclidean distance within their mutually possible class, and adding a corrective term depending on its location in the comparability graph. This would simplify the arguments in section \ref{section:topology}, and could potentially find use in algorithmic implementation.

%\subsubsection*{References}
\begin{thebibliography}{20}

\bibitem{sklar_dirichlet}Sklar, M. (2014). Fast MLE computation for the Dirichlet multinomial. arXiv preprint arXiv:1405.0099.
\bibitem{sklar_bias}Sklar, M. (2022). Sampling Bias Correction for Supervised Machine Learning: A Bayesian Inference Approach with Practical Applications. arXiv preprint arXiv:2203.06239.
\bibitem{mendelson}Mendelson, B. (1990). Introduction to topology. Courier Corporation.
\bibitem{bradley}Bradley, T. D., Bryson, T., \& Terilla, J. (2020). Topology: A Categorical Approach. MIT Press.
\bibitem{lyon}Lyon, A. (2016). Kolmogorovs Axiomatisation and its Discontents. The Oxford handbook of probability and philosophy, 155-166.
\bibitem{hajek}Hjek, A. (2003). What conditional probability could not be. Synthese, 137(3), 273-323.
\bibitem{censi}Censi, A., Frazzoli, E., Lorand, J., \& Zardini, G. (2022). Categorification of Negative Information using Enrichment. arXiv preprint arXiv:2207.13589.
\bibitem{ieee}Kahan, W. (1996). IEEE standard 754 for binary floating-point arithmetic. Lecture Notes on the Status of IEEE, 754(94720-1776), 11.
\bibitem{kolmogorov}A. N. Kolmogorov. Foundations of the Theory of Probability. Chelsea Publishing Company, New York
(1956). 
\bibitem{heinemann}Heinemann, F. (1997). Relative Probabilities. Working paper, http://www. sfm. vwl. uni-muenchen. de/heinemann/publics/relative probabilities-intro. htm.
\bibitem{discrete}Matouek, J., \& Neetil, J. (2008). Invitation to discrete mathematics. OUP Oxford.

\end{thebibliography}

This document along with revisions is posted at github as https://github.com/maxsklar/relative-probability-finite-paper. See readme for contact information. Local Maximum Labs is an ongoing effort create an disseminate knowledge on intelligent computing.
\end{document}
